---
title: "Orvac - impacts of surveillance visit"
output:
  html_document:
    number_sections: yes
    self_contained: yes
    theme: united
    toc: yes
    toc_float: true
    toc_depth: 3
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = F}
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(truncnorm))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(orvacsim))
ggplot2::theme_set(theme_bw())


cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


plot_tte_hist <- function(m){
  
  par(mfrow = c(2, 2))
  
  med0 <- log(2)/m[, 1]
  med1 <- log(2)/m[, 2]
  
  hist(med0, probability = T, main = "")
  abline(v = 30, col = "red", lwd = 2)
  abline(v = median(med0), col = "blue", lwd = 2)
  hist(med1, probability = T, main = "")
  abline(v = 35, col = "red", lwd = 2)
  abline(v = median(med1), col = "blue", lwd = 2)
  hist(m[, 3], probability = T, main = "")
  abline(v = 35/30, col = "red", lwd = 2)
  abline(v = median(med1/med0) , col = "blue", lwd = 2)
  plot(c(0, 10), c(0, 10))
  legend(0, 5, legend=c("true med", "sample med"),
         col=c("red", "blue"), lty=1:2, cex=0.8)
  par(mfrow = c(1, 1))
}


tte_suff_stats <- function(obst, trt, cen, n){
  
  obst <- obst[1:n]
  trt <- trt[1:n]
  cen <- cen[1:n]
  
  n_uncen_0 <- sum(1-cen[trt == 0])
  n_uncen_1 <- sum(1-cen[trt == 1])
  
  tot_obst_0 <- sum(obst[trt == 0])
  tot_obst_1 <- sum(obst[trt == 1])
  
  
  return(list(n_uncen_0 = n_uncen_0,
              n_uncen_1 = n_uncen_1,
              tot_obst_0 = tot_obst_0,
              tot_obst_1 = tot_obst_1))
}

```

# Preamble

Further to the brief document on the surveillance visits in ORVAC, here we look at the impact this approach has on parameter estimates derived from the data that we collect.

# Data generating process

Is as before -- exponential distribution. Here we use rates such that the medians are 30 and 37 months in the control and treatment arms respectively.

```{r, echo = F, fig.cap="Figure 1. Underlying time to event data (solid vertical is true value, dashed is sample estimate)", fig.align="center", fig.width=7, fig.height=5}

dnames <- c("id", "trt", "accrt", "age", 
            "serot2", "serot3", "probt3", 
            "evtt", "fu1", "fu2", "cen", "obst")

cfg <- readRDS("../sim_01/tests/cfg-example.RDS")

cfg$b0tte <- log(2)/30
cfg$b1tte <- log(2)/37 - log(2)/30


set.seed(2)
d <- rcpp_dat(cfg)
d2 <- as.data.frame(copy(d))
colnames(d2) <- dnames

v0 <- d2$evtt[d2$trt == 0]
v1 <- d2$evtt[d2$trt == 1]

par(mfrow = c(1, 2))
hist(v0, probability = T, main = "Ctl")
abline(v = log(2)/cfg$b0tte, col = cbp[2], lwd = 2)
abline(v = median(v0), col = cbp[3], lwd = 2, lty = 3)
hist(v1, probability = T, main = "Trt")
abline(v = log(2)/(cfg$b0tte+cfg$b1tte), col = cbp[2], lwd = 2)
abline(v = median(v1), col = cbp[3], lwd = 2, lty = 3)
legend(0, 5, legend=c("true med", "sample med"),
       col=c(cbp[2], cbp[3]), lty=1:2, cex=0.8)
par(mfrow = c(1, 1))
```

# Censoring

As a simplistic starting point, here is the effect of censoring based solely on the maximum follow up age of 36 months and assuming that all participants were accrued simultaneously - this isn't what we implement in the ORVAC code, it is just useful as a simple starting point to get an idea of the changes to the distribution.

```{r, echo = F, fig.cap="Figure 2. Censoring at 36 months", fig.align="center", fig.width=5, fig.height=4}
look <- 30
d2 <- as.data.frame(copy(d)); colnames(d2) <- dnames
 
# 
d2$cen2 <- NA
d2$cen2 <- ifelse(d2$evtt + d2$age > cfg$max_age_fu_months &
                    as.numeric(rownames(d2)) <= cfg$looks[look], 1, d2$cen2)
d2$cen2 <- ifelse(d2$evtt + d2$age <= cfg$max_age_fu_months &
                    as.numeric(rownames(d2)) <= cfg$looks[look], 0, d2$cen2)
d2$obst2 <- NA
d2$obst2 <- ifelse(!is.na(d2$cen2) & d2$cen2 == 0, d2$evtt, d2$obst2)
d2$obst2 <- ifelse(!is.na(d2$cen2) & d2$cen2 == 1, 
                   pmin(cfg$max_age_fu_months - d2$age, d2$evtt ), d2$obst2)
hist(d2$obst2, main = "", xlab = "Time")
```
  
The ORVAC code has two implementations of censoring. The first is similar to the above, except that it also considers the time of the interim and time of accrual as well as age. The figure below compares the distribution of scale parameter estimates. 

The histogram in the figure below shows the posterior distribution of the scale parameter under no censoring. The kernel density is the posterior from censored data (using the first censoring implementation in the ORVAC code) and the vertical dashed line is the true parameter values used to generate the data. Only a single dataset is being summarised so you wouldn't expect these to line up exactly.
  
```{r, echo = F, fig.cap="Figure 3. Parameter posteriors from censored data", results = "hide", fig.align="center", fig.width=8, fig.height=4}  
cfg$use_alt_censoring <- 1

lsuff <- list()
lsuff$n_uncen_0 <- 485
lsuff$tot_obst_0 <- sum(d2$evtt[d2$trt == 0 & as.numeric(rownames(d2)) <= cfg$looks[look]])
lsuff$n_uncen_1 <- 485
lsuff$tot_obst_1 <- sum(d2$evtt[d2$trt == 1 & as.numeric(rownames(d2)) <= cfg$looks[look]])

m_nocen <- matrix(0, nrow = cfg$post_draw, ncol = 3)
m_nocen[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_0,  (1/cfg$prior_gamma_b) + lsuff$tot_obst_0)
m_nocen[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_1,  (1/cfg$prior_gamma_b) + lsuff$tot_obst_1)
m_nocen[,3] <- m_nocen[,1]/m_nocen[,2]

cfg$use_alt_censoring <- 0
lsuffstat <- rcpp_clin_set_obst(d, cfg, look)

m <- matrix(0, nrow = cfg$post_draw, ncol = 3)
rcpp_clin_interim_post(m, 
                       lsuffstat$n_uncen_0, lsuffstat$tot_obst_0,
                       lsuffstat$n_uncen_1, lsuffstat$tot_obst_1,
                       cfg$post_draw, cfg);

par(mfrow = c(1, 3))
hist(m_nocen[,1], prob = T, main = "control", xlim = c(0.01, 0.03), xlab = "mu0")
abline(v = cfg$b0tte, lwd = 2, lty = 3)
lines(density(m[,1]), col = cbp[2], lwd = 2)
hist(m_nocen[,2], prob = T, main = "treatment", xlim = c(0.01, 0.03), xlab = "mu1")
abline(v = cfg$b0tte + cfg$b1tte, lwd = 2, lty = 3)
lines(density(m[,2]), col = cbp[2], lwd = 2)
hist(m_nocen[,3], prob = T, main = "ratio", xlim = c(0.6, 1.8), xlab = "ratio")
abline(v = cfg$b0tte/(cfg$b0tte + cfg$b1tte), lwd = 2, lty = 3)
lines(density(m[,3]), col = cbp[2], lwd = 2)
par(mfrow = c(1, 1))
```

The trial design is such that participants will be followed up at approximately 14-21 and 28-55 days and thereafter every 6 months (approximately) from randomisation. This logic (including the stochastic aspects) is implemented in the second approach to censoring in the ORVAC code. The figure below compares the distribution of the parameter estimates under the second ORVAC visit-based censoring implementation. 

The histogram in the figure below shows the posterior distribution of the scale parameter under no censoring. The kernel density is the posterior from censored data (using the second censoring implementation in the ORVAC code that incorporates the visit regimes) and the vertical dashed line is the true parameter values used to generate the data. Again, only a single dataset is being summarised so you wouldn't expect these to line up exactly.

<!--
One of the implications of the visiting mechanism is that if the event occurred after the latest visit but before the interim, we won't have any knowledge of the event - there are other effects too. Overall, we have a systematic underestimate of the number of events at every interim and therefore our posteriors reflect this lower event rate as shown below which suggests a higher median time to event in each arm than is actually the case. The figures below give an insight into the kind of bias (for want of a better word) that might arise.
-->


```{r, echo = F, fig.cap="Figure 4. Parameter posteriors from censored data", results = "hide", fig.align="center", fig.width=8, fig.height=4}  
cfg$use_alt_censoring <- 0
lsuffstat <- rcpp_clin_set_obst(d, cfg, look)

m2 <- matrix(0, nrow = cfg$post_draw, ncol = 3)
rcpp_clin_interim_post(m2, 
                       lsuffstat$n_uncen_0, lsuffstat$tot_obst_0,
                       lsuffstat$n_uncen_1, lsuffstat$tot_obst_1,
                       cfg$post_draw, cfg);

par(mfrow = c(1, 3))
hist(m_nocen[,1], prob = T, main = "control", xlim = c(0.01, 0.03), xlab = "mu0")
abline(v = cfg$b0tte, lwd = 2, lty = 3)
lines(density(m2[,1]), col = cbp[2], lwd = 2)
hist(m_nocen[,2], prob = T, main = "treatment", xlim = c(0.01, 0.03), xlab = "mu1")
abline(v = cfg$b0tte + cfg$b1tte, lwd = 2, lty = 3)
lines(density(m2[,2]), col = cbp[2], lwd = 2)
hist(m_nocen[,3], prob = T, main = "ratio", xlim = c(0.6, 1.8), xlab = "ratio")
abline(v = cfg$b0tte/(cfg$b0tte + cfg$b1tte), lwd = 2, lty = 3)
lines(density(m2[,3]), col = cbp[2], lwd = 2)
par(mfrow = c(1, 1))
```

<!--
The estimates for the two arms look wildly different from the histograms that show the results from the simplistic censoring that was shown originally. However, the ratio of the two arms seems to be fairly closely aligned to the actual values. Additionally, this was just one simulated dataset, it could be a freak occurrence, but it is not. Below are the results when the data generation and analysis are run 1000 times and the mean value of the posterior computed - it shows the systematic nature of the bias and also the fact that the ratio estimate are robust.
-->

The above figures show that the different estimators are approximately aligned, but the repeat-nature of the estimators are used to assess bias.

The figure below shows the results from a simulation examining the repeat-nature of the estimators. The histograms show the distribution of the mean of the posterior under no censoring computed at each iteration of the simulation. The kernel density shows the distribution of the mean of the posterior under the ORVAC visit-based censoring and  the black dashed vertical lines show the true parameter values. The distributions now line up reasonably well with the true values, but the ORVAC censoring is more spread out than the uncensored version. We would expect this to be the case as the uncensored data contains more information and therefore the resulting posterior of the parameter estimates is more precise.


```{r, echo = F, results = "hide", fig.cap="Figure 5. Parameter posteriors from censored data simulated 1000 times", results = "hide", fig.align="center", fig.width=8, fig.height=4}

cfg$use_alt_censoring <- 0
nsim <- 1000
mr <- matrix(0, ncol = 6, nrow = nsim)

for(i in 1:nsim){
  
  d <- rcpp_dat(cfg)
  lsuffstat <- rcpp_clin_set_obst(d, cfg, look)
  d2 <- as.data.frame(copy(d)); colnames(d2) <- dnames
  
  # uncensored 
  lsuff <- list()
  lsuff$n_uncen_0 <- 470
  lsuff$tot_obst_0 <- sum(d2$evtt[d2$trt == 0 & as.numeric(rownames(d2)) <= cfg$looks[look]])
  lsuff$n_uncen_1 <- 470
  lsuff$tot_obst_1 <- sum(d2$evtt[d2$trt == 1 & as.numeric(rownames(d2)) <= cfg$looks[look]])
  
  # non-informative (improper) prior
  m_nocen <- matrix(0, nrow = cfg$post_draw, ncol = 3)
  m_nocen[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_0,  ( 1/cfg$prior_gamma_b) + lsuff$tot_obst_0)
  m_nocen[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_1, ( 1/cfg$prior_gamma_b) +  lsuff$tot_obst_1)
  m_nocen[,3] <- m_nocen[,1]/m_nocen[,2]
  
  post_ctl <- mean(m_nocen[,1])
  post_trt <- mean(m_nocen[,2])
  post_ratio <- mean(m_nocen[,3])
  
  # censored using visits
  m <- matrix(0, nrow = cfg$post_draw, ncol = 3)
  rcpp_clin_interim_post(m, 
                         lsuffstat$n_uncen_0, lsuffstat$tot_obst_0,
                         lsuffstat$n_uncen_1, lsuffstat$tot_obst_1,
                         cfg$post_draw, cfg);
  
  post_ctl2 <- mean(m[,1])
  post_trt2 <- mean(m[,2])
  post_ratio2 <- mean(m[,3])
  
  mr[i,] <- c(post_ctl, post_trt, post_ratio, post_ctl2, post_trt2, post_ratio2)
}


par(mfrow = c(1, 3))
hist(mr[,1], prob = T, main = "control", xlim = c(0.01, 0.03), xlab = "mu0")
abline(v = cfg$b0tte, lwd = 2, lty = 3)
lines(density(mr[,4]), col = cbp[2], lwd = 2)
hist(mr[,2], prob = T, main = "treatment", xlim = c(0.01, 0.03), xlab = "mu1")
abline(v = cfg$b0tte + cfg$b1tte, lwd = 2, lty = 3)
lines(density(mr[,5]), col = cbp[2], lwd = 2)
hist(mr[,3], prob = T, main = "ratio", xlim = c(0.6, 1.5), xlab = "ratio")
abline(v = cfg$b0tte/(cfg$b0tte + cfg$b1tte), lwd = 2, lty = 3)
lines(density(mr[,6]), col = cbp[2], lwd = 2)
par(mfrow = c(1, 1))

```

<!--
Given that the ratio is robust to the way that we observe the data, the superiority check in the simulations should be ok. This is because, when make a superiority decision at the interim, it is only based on the posterior distribution for the ratio of the parameters. However, for the futility checks, we must use the posterior estimate for the rate parameter from each of the arms in order to obtain the posterior predictive distributions. 
-->

# Posterior predictive distribution

We can generate the posterior predictive distribution of the scale parameter by sampling from the posterior distributions for the scale parameter in each arm and then using these draws to generate data under the exponential distribution. The predictive distributions serve as a means to simulate future data. In turn, the simulated data is used to assess how many trials would end up successful if we ran the current trial to its maximum sample size.

From a single dataset, an uncensored posterior the proportion of successful trials was about 65%. When we added in the ORVAC visit-based censoring, the proportion of successful trials was above 42%. 

<!--
Simulation comparing the results derived from simplistic censoring versus the ORVAC visit-based censoring suggests that a much larger number of trials would designated as futile under the ORVAC censoring. In a comparison of a single dataset with median tte of 30 versus 40 months, around 78% of trials were designated as successful under the simple censoring and only 49% of the trials designated as successful under the ORVAC censoring. The difference in the posterior distributions for the ratio of the exponential rates (obtained after using the posterior predictive distributions to simulate data) are shown below (ORVAC is in grey, simple is yellow dashes).
-->

```{r, echo = F}
set.seed(2)
d <- rcpp_dat(cfg)
d2 <- as.data.frame(copy(d)); colnames(d2) <- dnames

lsuff <- list()
lsuff$n_uncen_0 <- 470
lsuff$tot_obst_0 <- sum(d2$evtt[d2$trt == 0 & as.numeric(rownames(d2)) <= cfg$looks[look]])
lsuff$n_uncen_1 <- 470
lsuff$tot_obst_1 <- sum(d2$evtt[d2$trt == 1 & as.numeric(rownames(d2)) <= cfg$looks[look]])

m_nocen <- matrix(0, nrow = cfg$post_draw, ncol = 3)
m_nocen[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_0,   ( 1/cfg$prior_gamma_b)  + lsuff$tot_obst_0)
m_nocen[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_1,  ( 1/cfg$prior_gamma_b)  + lsuff$tot_obst_1)
m_nocen[,3] <- m_nocen[,1]/m_nocen[,2]

# use posterior to derive pp
impute <- (1000 - cfg$looks[look])/2

# simulate 1000 trials
win <- numeric(1000)
mrat1 <- matrix(0, ncol = 1000, nrow = 1000)
mevt1 <- matrix(0, ncol = 1000, nrow = 1000)
matsuff1 <- matrix(0, ncol = 4, nrow = 1000)

for(i in 1:1000){
  
  pp <- matrix(0, nrow = impute, ncol = 2)
  pp[, 1] <- rexp(impute, m_nocen[i, 1])
  pp[, 2] <- rexp(impute, m_nocen[i, 2])
  
  # splice altenating together
  evtt_rep <- c(d2$evtt[1:cfg$looks[look]],   c(rbind(pp[, 2], pp[, 1])))
  mevt1[, i] <- evtt_rep
  
  lsuff$n_uncen_0 <- 500
  lsuff$tot_obst_0 <- sum(evtt_rep[d2$trt == 0])
  lsuff$n_uncen_1 <- 500
  lsuff$tot_obst_1 <- sum(evtt_rep[d2$trt == 1])
  
  matsuff1[i, ] <- c(lsuff$n_uncen_0, lsuff$n_uncen_1, 
                     lsuff$tot_obst_0, lsuff$tot_obst_1)

  m_rep <- matrix(0, nrow = cfg$post_draw, ncol = 3)
  m_rep[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_0, ( 1/cfg$prior_gamma_b) + lsuff$tot_obst_0)
  m_rep[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuff$n_uncen_1, ( 1/cfg$prior_gamma_b) + lsuff$tot_obst_1)

  mrat1[, i] <- m_rep[, 1] / m_rep[, 2]
  
  win[i] <- ifelse(mean(m_rep[, 1] / m_rep[, 2] > 1) > 0.975, 1, 0)
  
}
mean(win)

```



```{r, echo = F}
set.seed(2)
d <- rcpp_dat(cfg)

lsuffstat <- rcpp_clin_set_obst(d, cfg, look)
d2 <- as.data.frame(copy(d)); colnames(d2) <- dnames

m <- matrix(0, nrow = cfg$post_draw, ncol = 3)
rcpp_clin_interim_post(m,
                         lsuffstat$n_uncen_0, lsuffstat$tot_obst_0,
                         lsuffstat$n_uncen_1, lsuffstat$tot_obst_1,
                         cfg$post_draw, cfg);

# m <- matrix(0, nrow = cfg$post_draw, ncol = 3)
# m[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuffstat$n_uncen_0,   ( 1/cfg$prior_gamma_b)  + lsuffstat$tot_obst_0)
# m[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuffstat$n_uncen_1,  ( 1/cfg$prior_gamma_b)  + lsuffstat$tot_obst_1)
# m[,3] <- m[,1]/m[,2]


# use posterior to derive pp
impute <- (1000 - cfg$looks[look])/2

mevt2 <- matrix(0, ncol = 1000, nrow = 1000)
mrat2 <- matrix(0, ncol = 1000, nrow = 1000)
matsuff2 <- matrix(0, ncol = 4, nrow = 1000)

# simulate 1000 trials
win <- numeric(1000)
for(i in 1:1000){
  
  d3 <- as.data.frame(copy(d)); colnames(d3) <- dnames
  
  pp <- matrix(0, nrow = impute, ncol = 2)
  pp[, 1] <- rexp(impute, m[i, 1])
  pp[, 2] <- rexp(impute, m[i, 2])
  
  # splice altenating together
  d3$evtt <- c(d3$evtt[1:cfg$looks[look]], c(rbind(pp[, 2], pp[, 1])))
  
  mevt2[, i] <- d3$evtt
  
  lsuffstat <- rcpp_clin_set_obst(as.matrix(d3), cfg, length(cfg$looks))
  #lsuffstat
  matsuff2[i, ] <- c(lsuffstat$n_uncen_0, lsuffstat$n_uncen_1, 
                     lsuffstat$tot_obst_0, lsuffstat$tot_obst_1)
  
  m_rep <- matrix(0, nrow = cfg$post_draw, ncol = 3)
  rcpp_clin_interim_post(m_rep,
                         lsuffstat$n_uncen_0, lsuffstat$tot_obst_0,
                         lsuffstat$n_uncen_1, lsuffstat$tot_obst_1,
                         cfg$post_draw, cfg);
  
  # m_rep <- matrix(0, nrow = cfg$post_draw, ncol = 3)
  # m_rep[,1] <- rgamma(1000, cfg$prior_gamma_a + lsuffstat$n_uncen_0, ( 1/cfg$prior_gamma_b) + lsuffstat$tot_obst_0)
  # m_rep[,2] <- rgamma(1000, cfg$prior_gamma_a + lsuffstat$n_uncen_1, ( 1/cfg$prior_gamma_b) + lsuffstat$tot_obst_1)
  
  mrat2[, i] <- m_rep[, 1] / m_rep[, 2]
  
  win[i] <- ifelse(mean(m_rep[, 1] / m_rep[, 2] > 1) > 0.975, 1, 0)
  
}
mean(win)

```

The figures below show the posterior estimates for the ratio of scale parameters generated for 1000 simulations using the posterior predictive distribution. The distributions derived from the uncensored data is more precise than the ORVAC, but again, we would anticipate that this would be the case. 


```{r, echo = F, results = "hide", fig.cap="Figure 6. Ratios estimated from simulated trials using the posterior predictive distributions for time to events", results = "hide", fig.align="center", fig.width=7, fig.height=5}

# dfig0 <- data.frame(Simple = colMeans(mrat1), 
#                     ORVAC = colMeans(mrat2))
# dfig0 <- dfig0 %>%
#   tidyr::gather(key = "censoring", value = "mu")

dfig1 <- as.data.frame(mrat1) %>%
  tidyr::gather(key = "sim", value = "mu")
dfig1$sim <- as.numeric(gsub("V", "", dfig1$sim))
dfig1$censoring <- "Uncensored"


dfig2 <- as.data.frame(mrat2) %>%
  tidyr::gather(key = "sim", value = "mu")
dfig2$sim <- as.numeric(gsub("V", "", dfig2$sim))
dfig2$censoring <- "ORVAC"


dfig <- rbind(dfig1, dfig2)

idx1 <- sample(1:100, 100, replace = F)

ggplot(dfig %>% dplyr::filter(sim %in% idx1) ) +
  geom_line(aes(x=mu, colour = censoring, linetype = censoring , 
                group = factor(paste0(sim, censoring))), 
            stat="density", size=0.5, alpha=0.5) + 
  # geom_density(data = dfig0,
  #           aes(x=mu, colour = censoring, weight=mu/sum(mu)), size=0.5) +
  scale_x_continuous("Estimate for ratio of posteriors") +
  scale_y_continuous("Probability density") +
  scale_color_manual(values = cbp) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```




<!--
Figure 7 shows the distribution of sufficient statistics under differing censoring mechanisms obtained through simumlating a single trial which included predictive check on futility. 



```{r, echo = F, results = "hide", fig.cap="Figure 7. Sufficient stats under differing censoring", results = "hide", fig.align="center", fig.width=7, fig.height=5}


dfig1 <- as.data.frame(matsuff1)
dfig1$censoring <- "Simple"
dfig2 <- as.data.frame(matsuff2)
dfig2$censoring <- "ORVAC"

dfig <-rbind(dfig1, dfig2)
colnames(dfig) <- c("nobs0", "nobs1", "totobst0", "totobst1", "censoring")

dfig <- dfig %>%
  tidyr::gather(key = "variable", value = "value", -censoring)

dfig$variable <- factor(dfig$variable, 
                        levels = c("nobs0", "nobs1", "totobst0", "totobst1"),
                        labels = c("Num events (ctl)", "Num events (trt)", 
                                   "Total time (ctl)", "Total time (trt)"))

ggplot(dfig) +
  geom_line(aes(x=value, colour = censoring, linetype = censoring , 
                group = censoring), 
            stat="density", size=0.5) + 
  scale_x_continuous("Distributions of sufficient stats") +
  scale_y_continuous("Probability density") +
  scale_color_manual(values = cbp) +
  facet_wrap(~variable, scales = "free") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))



```

-->


# Conclusion

The results look reasonable and I can proceed with the simulations again.



<!--
```{r, echoh = F, results = "hide", fig.cap="Figure 6. Simulated event data from posterior predictives simulated trials", results = "hide", fig.align="center", fig.width=5, fig.height=5}

dfig1 <- as.data.frame(mevt1) %>%
  tidyr::gather(key = "sim", value = "mu")
dfig1$sim <- as.numeric(gsub("V", "", dfig1$sim))
dfig1$censoring <- "Simple"

dfig2 <- as.data.frame(mevt2) %>%
  tidyr::gather(key = "sim", value = "mu")
dfig2$sim <- as.numeric(gsub("V", "", dfig2$sim))
dfig2$censoring <- "ORVAC"

dfig <- rbind(dfig1, dfig2)

idx1 <- sample(1:100, 100, replace = F)

ggplot(dfig %>% dplyr::filter(sim %in% idx1) ) +
  geom_line(aes(x=mu, colour = censoring, linetype = censoring , 
                group = factor(paste0(sim, censoring))), 
            stat="density", size=0.5, alpha=0.2) + 
  # geom_density(data = dfig0,
  #           aes(x=mu, colour = censoring, weight=mu/sum(mu)), size=0.5) +
  scale_x_continuous("Estimate for ratio of posteriors", lim = c(0, 200)) +
  scale_y_continuous("Probability density") +
  scale_color_manual(values = cbp)
```
-->

