---
title: "Simulations to explore accrual rate influence on power/ppos"
subtitle: "Accrual and Dichotomous endpoint"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(pwr))
suppressPackageStartupMessages(library(WebPower))

ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)

# logit to p
inv_logit <- function(x){
  return(exp(x)/(1+exp(x)))
}
# p to logit
logit <- function(p){
  return(log(p/(1-p)))
}
prob_to_odd <- function(x){
  return(x/(1-x))
}
odd_to_prob <- function(x){
  return(x/(1+x))
}

```


# Preamble

Assume you have interims starting at $n_{min}$ then every $n$ patients up to a maximum sample size $N$. If you vary $n$ per interim (the accrual rate) how does this effect power in a dichotomous endpoint modelled with logistic regression?

Standard sample size calcs for power associated with effect of $\delta = p_1 - p_0$.

```{r}
N <- 250
p0 <- 0.3
p1 <- 0.4
pwr::pwr.2p.test(pwr::ES.h(p1, p0), N/2, alternative = "greater")

wp.logistic(n = 250, p0, p1 , alpha = 0.05, 
            power = NULL, alternative = c("greater"),
            family = c("Bernoulli"))
```

Simulation based approach requires us to simulate data.

```{r}
trial_dat <- function(trialid, n, p0, delta){
  d <- data.frame(trialid = trialid, id = 1:N)
  d$trt <- rep(0:1, len = 250)
  d$y <- rbinom(N, 1, inv_logit(b0 + b1*d$trt))
  d
}

nsim <- 1000
N <- 250
p0 <- 0.3
delta <- 0.1
p1 <- p0 + delta

b0 <- log(p0/(1-p0))
b1 <- log((p1/(1-p1))/(p0/(1-p0)))

d <- lapply(1:nsim, trial_dat, N, p0, delta)

head(d[[1]])
```

Simulation method to analysing power (ignore interims).

```{r}

sim <- function(trialid){
  lm1 <- glm(y~trt, dat = d[[trialid]][1:N,], family = binomial)
  # 1 sided wald: 
  pnorm(summary(lm1)$coefficients["trt","z value"], lower.tail = F)
}
p <- unlist(lapply(1:nsim, sim))
mean(p < 0.05)
```

Incorporate interim analyses that happen at following sample sizes (dictated by accrual rate).

```{r}
get_int <- function(n){
  nmin <- 70
  interm_n <- seq(nmin, N, by = n)
  if(max(interm_n)!=N){
    interm_n <- c(interm_n, N)
  } 
  interm_n
}
get_int(n = 30)
```

Revised simulation.

```{r}

fit <- function(current_n, trial_id){
  lm1 <- glm(y~trt, data=d[[trial_id]][1:current_n,], family = binomial)
  # 1 sided wald: 
  pnorm(summary(lm1)$coefficients["trt","z value"], lower.tail = F)
}

sim <- function(trial_id, n_step){
  res <- unlist(lapply(get_int(n_step), fit, trial_id))
  any(p.adjust(res, method = "bonferroni") < 0.05)
}


# ppos
accrual_rates <- c(25, 30, 35, 40, 45, 50)
ppos <- rep(0, length(accrual_rates))

d <- lapply(1:nsim, trial_dat, N, 0.4, 0.15)
head(d[[1]])

for(i in 1:length(accrual_rates)){
  ppos[i] <- mean(unlist(lapply(1:nsim, sim, accrual_rates[i])))
}

```

Plot.

```{r}
plot(accrual_rates, ppos, type = "l", ylim = c(0.2, 0.5))
```

# Summary

For the given samplesize $N$ with $n$ participants being analysed per interim, PPOS gradually increases as accrual rate increases.

