---
title: "Effects of baseline seroconversion rate on power"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(pwr))
suppressPackageStartupMessages(library(WebPower))

ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)

# logit to p
inv_logit <- function(x){
  return(exp(x)/(1+exp(x)))
}
# p to logit
logit <- function(p){
  return(log(p/(1-p)))
}
prob_to_odd <- function(x){
  return(x/(1-x))
}
odd_to_prob <- function(x){
  return(x/(1+x))
}

```

# Preamble

For given observations $N$ and a given difference in probability of events in each arm, is logistic regression more powerful when the extreme values are near 0 and 1?

Generate data.

```{r}
N <- 250

p0 <- 0.3
delta <- 0.1
p1 <- p0 + delta

b0 <- log(p0/(1-p0))
b1 <- log((p1/(1-p1))/(p0/(1-p0)))

gen_dat <- function(){
  d <- data.frame(id = 1:N)
  d$trt <- rep(0:1, len = 250)
  d$y <- rbinom(N, 1, inv_logit(b0 + b1*d$trt))
  d
}
```

Standard sample size calcs for power associated with effect of $\delta=p_1−p_0$.

```{r}
pwr::pwr.2p.test(pwr::ES.h(p1, p0), N/2, alternative = "greater")
```

```{r}
wp.logistic(n = 250, p0, p1 , alpha = 0.05, 
            power = NULL, alternative = c("greater"),
            family = c("Bernoulli"))
```

Simulation based approach requires data generation.

```{r}
nsim <- 1000
sim <- function(){
  lm1 <- glm(y~trt, dat = gen_dat(), family = binomial)
  # 1 sided wald: 
  pnorm(summary(lm1)$coefficients["trt","z value"], lower.tail = F)
}
p <- replicate(nsim, sim())
mean(p)
```

Vary the baseline.

```{r}
p0 <- seq(0.05, 0.85, by = 0.05)
p1 <- p0 + delta
p <- numeric(length(p0))

for(i in 1:length(p0)){
  wp <- wp.logistic(n = 250, p0[i], p1[i], alpha = 0.05, 
            power = NULL, alternative = c("greater"),
            family = c("Bernoulli"))
  p[i] <- wp$power
}
```

Plot power versus baseline.

```{r}
plot(p0, p, type = "l", ylim = c(0, 1))
```



# Summary

For a given absolute difference $\delta = p_1 - p_0$ the power of detecting a difference using logistic regression varies as a function of the baseline probability.

